---
title: "L06: Samples and Observational studies"
date: "July 13, 2020"
---

<!-- libraries -->
```{r,include=FALSE,purl=FALSE}
library(knitr) # for include_graphics() 
library(dplyr)
```

<!-- define default size for R graphics -->
```{r,include=FALSE,purl=FALSE}
outputFormat<-opts_knit$get('rmarkdown.pandoc.to')
if(outputFormat=="beamer"){
  opts_chunk$set(fig.width=6,fig.height=4)
}
```

<!-- define function for external images -->
```{r,include=FALSE,purl=FALSE}
image<-function(ff,ss,cc=NULL,ll=NULL){
  if(is.null(cc)){
    paste('\\centering','\n',
          '\\includegraphics[scale=',ss,']{',ff,'}',
          sep='')
  } else {
    paste('\\centering','\n',
          '\\copyrightbox[b]{',
          '\\includegraphics[scale=',ss,']{',ff,'}',
          '}{\\raggedleft{\\tiny \\href{',ll,'}{',cc,'}}}',
          sep='')    
  }
}
```

<!-- define function for links -->
```{r,include=FALSE,purl=FALSE}
link<-function(tt,ll){
  paste('[\\textcolor[HTML]{ffa328}{\\ul{',tt,'}}]','(',ll,')',sep='')
}
```


## Sampling in R

### Example of SRS in R

- First read in the hospital cesarean data

```{r, message=F, echo=FALSE}
library(readxl)
library(dplyr)
CS_data <- read_xlsx("kozhimannil.xlsx", sheet = 1)
CS_data <- CS_data %>% mutate(ID = row_number())
head(CS_data)
```


### Example of SRS in R

```{r sample-n-100}
CS_100_1 <- CS_data %>% sample_n(100)
CS_100_2 <- CS_data %>% sample_n(100)
```

Do you expect `head(CS_100_1)` to equal `head(CS_100_2)`?

### Example of SRS in R

```{r are-first-rows-the-same}
head(CS_100_1 %>% select(Births, HOSP_BEDSIZE, cesarean_rate, ID))
```

### Example of SRS in R
```{r are-first-rows-the-same2}
head(CS_100_2 %>% select(Births, HOSP_BEDSIZE, cesarean_rate, ID))
```

### Example of SRS in R
```{r are-first-rows-the-same3}
identical(CS_100_1, CS_100_2)
```

### Example of SRS in R

Why are these first six lines different?

Anytime you do something *randomly* in R, the results will be different. This is
a good thing! This allows you to pick many different random samples. In future
weeks we will do this a lot.


### Example of SRS in R

What if you want to ensure that you pick the same SRS as a friend?

Then you need to use `set.seed()`:

```{r}
set.seed(123)
CS_100_1 <- CS_data %>% sample_n(100)

set.seed(123)
CS_100_2 <- CS_data %>% sample_n(100)

identical(CS_100_1, CS_100_2)
```

### SRS a fraction in R

```{r}
CS_5percent <- CS_data %>% sample_frac(0.05)
```

### Proportionate Stratified sampling in R

- Group *and then* sample the same fraction from each group

```{r}
CS_10percent_grouped <- CS_data %>% 
  group_by(HOSP_BEDSIZE) %>%
  sample_frac(0.1)
dim(CS_10percent_grouped)
```

- Proportionate stratified SRS assembles a sample that maintains the relative 
proportions of `HOSP_BEDSIZE` in the chosen sample compared to the population

### Proportionate Stratified sampling in R

How to check you really did sample 10% of each `HOSP_BEDSIZE` group?

First see how many hospitals fall into each category in the original data
```{r ,out.width="80%"}
CS_data %>% group_by(HOSP_BEDSIZE) %>% tally()
```

### Proportionate Stratified sampling in R
then in the sample
```{r ,out.width="80%"}
CS_10percent_grouped%>%group_by(HOSP_BEDSIZE) %>% tally()
```
