---
title: "Homework 2: Summarizing COVID-19 Pandemic Data from US Counties"
author: "Your name and student ID"
date: "Today's date"
output: pdf_document
---

* Due date: Monday, July 13 10:00pm.
* Remember: autograder is meant as sanity check ONLY. It will not tell you if you have the correct answer. It will tell you if you are in the ball park of the answer so *CHECK YOUR WORK*.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/hw02.RAGS.R")
```

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage

--------------------------------------------------------------------------------

# Summarizing COVID-19 Pandemic Data from US Counties

### Introduction 

In this week's assignment, you will describe COVID-19 pandemic data
distributions using numbers. The data is current to 6/29/2020 and centered
on the county level. You will investigate the **mean** and **median** of different 
attributes and also examine the distribution of COVID-19 data based on different factors.
Lastly, you will describe the **spread**
of the distributions using **quartiles** and make a **box plot**. 

Execute this code chunk to load the required libraries:
```{r load-libraries}
library(readr)
library(dplyr)
library(ggplot2)
```

Just like in homework 1, read in the data that is stored as a .csv file and assign 
it to an object called `covid_data`. We will also use dplyr's `mutate()` to create
the a new dataframe column with the uninsured rate that ranges between 0 and 100: 

```{r read-in-the-data}
covid_data <- read_csv("us-county-covid.csv")

covid_data <- covid_data %>% mutate(uninsured_rate = num_uninsured/population)
covid_data
```

\newpage

**1. [1.5 points] Fill in the blanks indicated by "----" by saving the answer to each blank in the code chunk below. Make sure you capitalize correctly, as R is case-sensitive.**

The function `mutate()` takes the old variable called **--(a)--** and divides it
by **--(b)-** to make a new variable called **--(c)--**. 

Note: Make sure you leave your answer is quotes! You are assigning object a, b, c to a word (not a number, function, or other object) so leave you answer in quotes.

```{r p1}
a <- "<<<<YOUR ANSWER HERE>>>>"
b <- "<<<<YOUR ANSWER HERE>>>>"
c <- "<<<<YOUR ANSWER HERE>>>>"

# ------- REMINDER -------
# The checks on this homework are only provided as sanity checks.
# They do not guarantee correctness.
# ------------------------
check_problem1()
```

\newpage

**Investigate what would have happened had we not assigned the data using `<-` to 
`covid_data`? Re-run the code without the assignment and see what happens.**

```{r no-assignment-example}
# First, let's re-read in the data as we did in the previous chunk
covid_data <- read_csv("us-county-covid.csv")

# Now, we try again without the re-assignment to covid_data
covid_data %>% mutate(uninsured_rate = num_uninsured/population)

# check the variables on covid_data
names(covid_data)
```

Did `uninsured_rate` get added to the data set? No. You can tell by using 
`names(covid_data)` to view the names of your columns and notice that the variable hasn't been added. This is because when we don't assign the output to anything, 
it just prints it out for us to see. Nothing is saved. So, we want to save the 
output by assigning the result of the code to a variable, which in this case, we
used `covid_data`. In general, you want to use **new variable names** at every 
significant step in your analysis as you work with your data, so that you have 
access to the data at all those significant stages. However, if you are performing 
multiple small operations on the same dataset, you can overwrite the original 
variable, since you know you won't be needing the in-between steps anyway.

```{r overwrite-original-data}
# This overwrites the original covid_data object
covid_data <- covid_data %>% mutate(uninsured_rate = num_uninsured/population)
```

\newpage

### GDP: Summarizing measures of centrality

Check out this histogram!
```{r histogram-county california}
ggplot(data = covid_data, aes(x = uninsured_rate)) +
  geom_histogram(col = "white", fill = "sienna2") + 
  xlab("uninsured_rate") +
  theme_minimal() 
```

**2. [1 point] Describe the shape of this distribution. Is it "skewed left", "skewed right", "symmetric", or "bimodal"? Uncomment one of the possible choices.**
```{r histogram-GDP-2006-modal}
# p2 <- "skewed left"
# p2 <- "skewed right"
# p2 <- "symmetric"
# p2 <- "bimodal"

check_problem2()
```
\newpage

**3. [1 point] Based on your answer, will the mean be approximately the "same", "larger than", or "smaller than" the median?**

```{r histogram-GDP-2006}
# p3 <- "same"
# p3 <- "larger than"
# p3 <- "smaller than"

# This only checks that you've selected an answer, not its correctness.
check_problem3()
```

\newpage

**4. [3 points] Describe, in words, how the mean and median are calculated:**

[TODO: YOUR ANSWER HERE]

\newpage
To calculate the mean and median in R, we can use the `summarize()` function 
from the `dplyr` package. The `summarize()` function is used anytime we want to 
take a variable and summarize something about it into one number, like it's mean 
or median. Here is the code to summarize `uninsured_rate`'s mean and print it out to 
the screen. In the code, we name the mean `mean_uninsured` and output the result to 
the screen:

```{r summarize-mean}
uninsured_summary <- covid_data %>% summarize(mean_uninsured = mean(uninsured_rate))
uninsured_summary
```
\newpage

**5. [1 point] Extend the above code to also summarize the median. Make sure to include both the mean and the median in covid_summary**
Call the median summary `median_uninsured`. Assign this summary to `covid_summary`

```{r summarize-median}
covid_summary <- "<<<<YOUR CODE HERE>>>>"
covid_summary

check_problem5()
```
\newpage

**6. [2 points]** `geom_vline()` can be used to add the mean and the median to the 
histogram shown above. This `geom` adds a vertical line to the graph. You need 
to specify where to add the line by passing it an "x-intercept" argument. Remove 
the comments (i.e., the three "#") from the code below and update the 
`geom_vline()` code to plot lines at the mean and median by telling it the mean 
and median estimates. Round your values to 3 decimal places. This time, use the mean and median of the `uninsured_rate`, instead of `num_uninsured`. 

The argument `lty=1` (standing for line type) will plot a solid line and `lty=2` will plot a dashed line.  

**For the purposes of this question, please assign xintercept to a plain NUMERIC, not a variable or expression**

```{r add-mean-med-plot}

p6 <- ggplot(data = covid_data, aes(x = uninsured_rate)) +
  geom_histogram(col = "white", fill = "sienna2") + 
  xlab("Uninsured Rate") +
  theme_minimal()  #+
        #geom_vline(aes(xintercept = ), lty=1) + 
        #geom_vline(aes(xintercept = ), lty=2) 
p6


# REMINDER: this is only a sanity check; it doesn't check for accuracy
check_problem6()

```

\newpage

### Summarizing the deaths by state

Here's the distribution of age_adjusted_deaths_rates across counties:

```{r CS-histogram}
ggplot(data = covid_data, aes(x = age_adjusted_death_rate)) +
  geom_histogram(binwidth = 4, col = "white", fill = "sienna2") + 
  xlab("age_adjusted_death_rate") +
  theme_minimal()
```
Note: You may have noticed a warning that stated "Removed 55 rows containing non-finite values (stat_bin)." All that tells you is that some rows did not have a value for this variable, instead they were 



\newpage

**7. [1 point] Describe the shape of this distribution. Is it "skewed left",
"skewed right", "symmetric", or "bimodal"?**

```{r histogram-GDP-2006-distribution}
# p7 <- "skewed left"
# p7 <- "skewed right"
# p7 <- "symmetric"
# p7 <- "bimodal"

# This only checks that you've selected an answer, not its correctness.
check_problem7()
```

\newpage

**8. [1 point] Extend the ggplot code given below using the `facet_wrap()` statement to make a separate histogram for each state of the `uninsured_rate` variable:**

```{r CS-histogram-by-income}
p8 <- ggplot(data = covid_data, aes(x = uninsured_rate)) +
        geom_histogram(col = "white", fill = "sienna2") + 
        xlab("uninsured rate (%)") +
        theme_minimal() #+ 
        #facet_wrap(YOUR CODE HERE) 
p8


check_problem8()
```
\newpage

**9. [2 points] Based on this plot describe the differences in uninsured rates between California and Texas**

[TODO: YOUR ANSWER HERE]



\newpage

**10. [1 point] How might uninsured rates affect COVID-19 mortality rates?**

[TODO: YOUR ANSWER HERE]




\newpage

**11. [2 points] Calculate the `mean_uninsured_ca` and `median_uninsured_ca` of `uninsured_rate` in California's counties using only the `summarize()` command and a dplyr function. Assign this summary to the name `uninsured_summary` and then 
print the results by typing `uninsured_summary` so you can see the contents.**

```{r calc-CS-mean-median}
uninsured_summary <- "<<<<YOUR CODE HERE>>>>"

uninsured_summary

check_problem11()
```
\newpage

**Measures of variation**

**12. [2 marks] Use ggplot2 to make a boxplot of the distribution of `uninsured_rate`**

```{r}
p12 <- "<<<<YOUR CODE HERE>>>>"
p12


check_problem12()
```
\newpage

Recall that the box plot summarizes the distribution in five numbers: the minimum, 
the first quartile (with 25% of the data below it), the median, the third quartile 
(with 75% of the data below it), and the maximum. Each of these numbers has at 
least one corresponding R function:

| Number          | R function                                                   |
|-----------------|--------------------------------------------------------------|
| Minimum         | `min(variable)`                                              |
| First quartile  | `quantile(variable, probs = 0.25)`                           |
| Median          | `median(variable)` or `quantile(variable, probs = 0.5)`      |
| Third quartile  | `quantile(variable, probs = 0.75)`                           |
| Maximum         | `max(variable)`                                              |

**13. [2 points] Use a combination of `dplyr`'s `summarize` function and the 
above functions to compute the five number summary of `uninsured_rate`. Assign the 
summary to the name `five_num_summary`, which should contain values for `min`, `Q1`,
`median`, `Q3`, and `max`.**

```{r five-number-summary}
five_num_summary <- "<<<<YOUR CODE HERE>>>>"
five_num_summary


check_problem13()
```
\newpage

Double check that `geom_boxplot()` is making the box plot correctly. You can do 
this by adding horizontal lines to the plot at each number in your five number 
summary using `geom_hline()`. Because horizontal lines intercept the y-axis, 
`geom_hline()` requires the `yintercept` argument that you can set to each number 
in your summary. 

**14. [2 points] The code below includes one horizontal line at the minimum shown in blue. Add 
the rest of the lines using geom_hline. Round to 2 decimal places and up if needed:**

```{r add-horizontal-lines-to-boxplot}
p14 <- ggplot(data = covid_data, aes(y = uninsured_rate)) +
          geom_boxplot(col = "black", fill = "sienna2") +
          theme_minimal() + 
          geom_hline(aes(yintercept = 0.06), col = "blue")
p14


check_problem14()
```

\newpage

15. [4 points] **BONUS** At the end of every homework, you will have the opportunity to obtain up to *4 extra credit points* on that assignment. 

+1 point for developing a multiple choice question on the material in the homework

+1 point for developing a true/false question on the material 

+2 points for developing a free response question.

To get full credit, you need to write the question, write an answer, provide a rubric for grading the answer, and say something about why/how you feel this question addresses a concept in this homework. You can only write **one question per category**.

*Note: If you are including code as part of your questions, please include an additional code chunk that loads the required packages and the required dataset.*

[ YOUR ANSWER HERE ]

### Check your score



Click on the middle icon on the top right of this code chunk (with the downwards gray arrows and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su20/hw/hw02; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.
