---
title: "Lab 2: Visualization of global rates of cesarean delivery"
author: "Your name and student ID"
date: "today's date"
output: pdf_document
---

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab02.RAGS.R")
```

### Instructions 
* Due date: Monday, July 13 at 10:00pm.
* Late penalty: 50% late penalty if submitted within 24 hours of due date, no 
marks for assignments submitted thereafter.
* This assignment is graded on **correct completion**, all or nothing. You must pass all public tests and submit the assignment for credit.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

### Visualizing global cesarean delivery rates and GDP across 137 countries

The proportion of pregnant women who undergo cesarean deliveries varies 
dramatically across countries. In the US, the cesarean delivery rate is 
around 33%. A 2018 study in the British Medical Journal found that cesarean 
delivery rates ranged between 0.6% in South Sudan and 58.9% in the Dominican 
Republic.^1^

Over the course of this lab and the first assignment, we will study the relationship 
between countries' cesarean delivery rates and their gross domestic products (GDPs).^2,3^ 

In today's lab we will:

- import the data,
- familiarize ourselves with the dataset, 
- explore the univariate distributions of the key variables.

If you haven't already, begin by knitting this document by clicking the "Knit" 
button above. As you fill it out, you can re-knit (clicking the button again) and 
see how the document changes. 

Let's get started!

Click the green arrow icon in the top right corner of the gray chunk below to execute 
the three lines of code in the code chunk, or execute them line by line by placing your 
cursor on the first line and hitting cmd + enter on Mac or ctrl + enter on PC. If you 
do not want messages produced by loading `dplyr` to print in your knitted document, 
toggle `message = TRUE` to `message = FALSE` in the code chunk options. Re-knit your file
and notice the different in the output:

```{r load-libraries, message = TRUE}
library(readr)
library(dplyr)
library(ggplot2)
```

- The `library` command loads the packages into memory. 
- The `readr` library contains function for reading in data.
- The `dplyr` library contains functions we will use to manipulate data.
- The `ggplot2` library contains functions to visualize data.


The data on each country's cesarean delivery rate is stored in a comma-separated-values (csv) file. We can load that data and assign it to an object called
`CS_data` by running the code below. Click the green arrow in the line below to 
run the code. (Note that the green arrow will run all the code in this code chunk,
not just the first line of code.):

*To access tutorial for this section, follow this link to Video 1 https://bcourses.berkeley.edu/courses/1493244/pages/lab-2*

```{r read-in-the-data}
CS_data_raw <- read_csv("cesarean.csv")

# This code re-orders the variable Income_Group in the specified order.
# Note that it *does not* change the order of the data frame (like arrange() does)
# Rather, it specifies the order the data will be plotted. 
# This will make more sense when we plot the data using Income_Group, and then
# again using Income_Group_order
CS_data_raw <- CS_data_raw %>% 
  mutate(Income_Group_order = forcats::fct_relevel(Income_Group, "Low income", 
                                              "Lower middle income", "Upper middle income", 
                                              "High income: nonOECD", "High income: OECD"))
```

Look over at the environment tab (on the top right pane) and make sure you see 
your dataset there. Click on `CS_data` in that pane. This will open the 
data viewer tab. Take a quick look at the data and then switch back to the 
"lab02-cesarean-delivery.Rmd" file and move to the next step.

Often, when you receive a dataset, you also receive a data dictionary that tells
you what each variable means. Below, is an example of such a dictionary for this 
dataset:

| Variable        | Description                                         |
|-----------------|-----------------------------------------------------|
| Country_Name    | The name of the country                             | 
| CountryCode     | 3-digit unique country identifier                   |                
| Births_Per_1000 | The number of births per 1000 people in the country |
| Income_Group     | The income group the country is categorized into    |
| Region          | The geographic region the country belongs to        |
| GDP_2006        | The gross domestic product (GDP) of the country in 2006 |
| CS_rate         | The percent of births that were delivered by cesarean section |
| Income_Group_order | The "ordered" version of Income_Group |

\newpage 

### Look at your newly imported data

In four separate code chunks, type `str(CS_data_raw)`, `dim(CS_data_raw)`, 
`names(CS_data_raw)`, and `head(CS_data_raw)`. Execute each code chunk and study the output.

```{r keep-exploring-the-data-1}
# <<<<YOUR CODE HERE>>>>

```

```{r keep-exploring-the-data-2}
# <<<<YOUR CODE HERE>>>>

```

```{r keep-exploring-the-data-3}
# <<<<YOUR CODE HERE>>>>

```

```{r keep-exploring-the-data-4}
# <<<<YOUR CODE HERE>>>>

```

**In your own words, what do these functions do:**

- `dim()`: [ TODO: YOUR ANSWER HERE ] 

- `str()`: [ TODO: YOUR ANSWER HERE ]


- `names()`: [ TODO: YOUR ANSWER HERE ]


- `head()`: [ TODO: YOUR ANSWER HERE ]


**Based on these functions' outputs, who/what are the individuals in this 
dataset? How many of them are there?**

[ TODO: YOUR ANSWER HERE ]



\newpage 

### Add a new variable

**1. Right now, `CS_rate` is presented as a number between 0 and 1. For plotting, it 
will be nice to have the proportion displayed as a number between 0 and 100. We 
can use `dplyr`'s `mutate()` function to add a new variable called `CS_rate_100`
to the dataset that takes this desired format. Write one line of code to add this
variable to the dataset and assign this to the new object `CS_data`:**

```{r add-new-variable}
CS_data <- "<<<<YOUR CODE HERE>>>>"

check_problem1()
```

### Examine the distribution of income and geographic regions

**These countries represent a wide range of income and geographic regions. By geographic regions?**

**Which variable groups the countries by income regions?** [ TODO: YOUR ANSWER HERE ]


**Which variable groups the countries by geographic regions?** [ TODO: YOUR ANSWER HERE ]


*To access tutorial for this section, follow this link to Video 2 https://bcourses.berkeley.edu/courses/1493244/pages/lab-2*


These variables are categorical variables. So far in lecture, we've seen categorical variables stored as `chr` variables when we've examined `str(data)`. Here however, `Income_Group_order` is stored as a `Factor` variable, which is another way to store categorical data. One benefit of storing it as `Factor` is that we can re-order the `Factor` variable in a way that makes sense for plotting and later analysis. To see the values and order of a factor variable, we can use the following code.

```{r levels-income}
CS_data %>% pull(Income_Group_order) %>% levels
# You won't be tested on writing the line of code above. 
# Try and gain an understanding of what it is showing you
```
\newpage 

**2. Because `Income_Group` and `Region` are categorical, we make bar charts to represent 
their distributions. Using the `ggplot()` code you learnt during lecture, make a 
bar chart showing the number of countries per income group using `Income_Group`.**

```{r bar-chart-income}
p2 <- "<<<<YOUR CODE HERE>>>>"
p2

check_problem2()
```

\newpage 

**3. Now, make the plot using `Income_Group_order`.**

```{r bar-chart-income-order}
p3 <- "<<<<YOUR CODE HERE>>>>"
p3

check_problem3()
```

**Do you prefer the plot using `Income_Group` or `Income_Group_order`? Why?**

[ TODO: YOUR ANSWER HERE ]



**Did you use the argument `stat="identity"` in the code to make this bar chart? Why or why not?**

[ TODO: YOUR ANSWER HERE ]



\newpage 

**4. Make a bar chart showing the number of countries per geographic region.**

```{r bar-chart-region, fig.width = 11}
p4 <- "<<<<YOUR CODE HERE"
p4

check_problem4()
```

**Based on your plot, which world region has the most countries in the data set?**

[ TODO: YOUR ANSWER HERE ]



**Which geographic region had the fewest number of countries in the data set**

[ TODO: YOUR ANSWER HERE ]



**Note that the above code chunk had the chunk option "fig.width = 11". Try changing the
setting to a lower number, like 9 and re-running the chunk. Notice how the width
of the rendered plot changes. Change it back to 11 so that the x-axis labels 
don't look too cramped.**

### Examine the distribution of `GDP_2006` and `CS_rate_100` across countries 

*To access tutorial for this section, follow this link to Video 3https://bcourses.berkeley.edu/courses/1493244/pages/lab-2*

**5. `GDP_2006` and `CS_rate_100` are quantitative, meaning that we use histograms 
rather than bar charts to examine their distributions. Using code you learnt in 
class, make a histogram of `GDP_2006`. Pay attention to the message that R 
outputs when running your code. It is telling you that by default the data is 
grouped into 30 bins for plotting.**

```{r histogram-GDP-2006}
p5 <- "<<<<YOUR CODE HERE>>>>"
p5

check_problem5()
```

**6. Rather than letting R choose the number of `bins` we can set `binwidth = number`
within the geom function to choose the width of the bins. Try setting the 
binwidth to different amounts and see how it changes the look of the plot.**

**To choose a binwidth, look at the x-axis of the histogram you just made. Because
we are plotting GDP across a wide variety of countries, it might make sense to 
choose a bin width in the 1000s. Thus, the range of the variable matters when
choosing a good bin width.** 

```{r histogram-GDP-2006-binwidth}
p6 <- "<<<<YOUR CODE HERE>>>>"
p6

check_problem6()
```

\newpage 

**7. Next, make a histogram for cesarean delivery rates using the variable 
`CS_rate_100`. First run your code without binwidth set and then add in the 
binwidth argument to choose the size.** 

```{r histogram-CS}
p7 <- "<<<<YOUR CODE HERE>>>>"
p7

check_problem7()

```

\newpage 

### Polish up your plots

*To access tutorial for this section, follow this link to Video 4 https://bcourses.berkeley.edu/courses/1493244/pages/lab-2*

**Your plots look great so far, but let's get them ready to present at a meeting 
or in a report. To do this please update the plot and axes titles (`labs`), and 
change the `fill` or `col`or of the bars. Here is a [big list](http://sape.inf.usi.ch/quick-reference/ggplot2/colour) 
of colors that you can set col and fill equal to. For example col = "orchid1"
will change the color to a pinky-purple.**

**8. Below present your final plots with updated titles, labels, colors, and fills.To do this, extend the ggplot code you wrote above with the functions and arguments just described:**

```{r your-final-plots}
p8 <- "<<<<YOUR CODE HERE>>>>"
p8

check_problem8()
```

\newpage 

### Save your plots to separate files

Sometimes, you need to save plots as PNG or JPEG to include in a PowerPoint 
presentation or submit to a research journal for publication. The `ggsave()`
function is used to save plots as separate files.

Quickly read the documentation of the `ggsave()` function by executing this code
which opens a help window in the bottom right pane:

```{r what-is-ggsave}
?ggsave
```

**9. Try figuring out how to save a plot based on the documentation and running your 
attempts in the code chunk below. Don't worry about producing an error or saving your 
plot in the wrong place -- that is all part of coding and learning new functions.**

```{r saving-plot}
# <<<<YOUR CODE HERE>>>>

```


\newpage 

**When you are done, knit your R document one last time and ensure it looks good! 
Change anything you don't like. You can choose to knit to .docx or .pdf by 
selecting these options form the dropdown menu next to knit, in the case that 
you'd like to save the completed file onto your computer.** 

When you're happy with your work, please follow the instructions on the final page to submit.

```{r check-score}
# Just run this
total_score()
```

### References

1. Boatin AA, Schlotheuber A, Betran AP, Moller AB, Barros AJ, Boerma T, Torloni MR, Victora CG, Hosseinpoor AR. Within country inequalities in caesarean section rates: observational study of 72 low and middle income countries. BMJ. 2018;24(360):k55. DOI: 10.1136/bmj.k55.

2. Gibbons L, Belizán JM, Lauer JA, Betrán AP, Merialdi M, Althabe F. The global numbers and costs of additionally needed and unnecessary caesarean sections performed per year: overuse as a barrier to universal coverage. World Health Organization; 2010. Available at: http://www.who.int/healthsystems/topics/financing/healthreport/30C-sectioncosts.pdf. Accessed October 4, 2015.

3. The World Bank. GDP per capita (current US$). Washington, DC; Available at: http://data.worldbank.org/indicator/NY.GDP.PCAP.CD. Accessed Sep 30, 2015.

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su20/lab/lab02; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.

