---
title: "Homework 4"
author: "Your name and student ID"
date: "Today's date"
output: 
  pdf_document: default
---
* Due date: Monday, July 20 10:00pm
* Remember: autograder is meant as sanity check ONLY. It will not tell you if you have the correct answer. It will tell you if you are in the ball park of the answer so *CHECK YOUR WORK*.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/hw04.RAGS.R")
```

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage

--------------------------------------------------------------------------------

**[12 points] Part 1: Simulating COVID-19 estimates and sampling from an infinitely large population** 

The Center for Disease Control and Prevention (CDC) estimates that 3 in every
50 Americans have COVID-19 in the United States.

1. [3 points] Define a random variable for "COVID-19". Write down the probability model for the random variable. Round each percentage to two decimal places (e.g., 0.43224 
would be rounded to 43.22%). Is the sample space discrete or continuous?

[ TODO: Replace the text with your list of proportions and your explanation. ]

You might want to use the table template below to write out your probability 
model. If not, then delete it. *Knit now* to see how this table is rendered in your PDF.

| Tables   |      Are     |  Cool |
|----------|--------------|-------|
| yadi     | type stuff   | X     |
| yadi     | more stuff   | Y     |
| yada     | etc          | Z     |


\newpage

2. [2 points] Simulate data that equals 0 if there is no COVID-19 case and equals
1 if there is a COVID-19 case. Simulate this data for 200 tests at a local 
hospital. Be sure to use the risk of COVID-19 from part a). Assign your 
simulated output the name `sim_01`. Print your simulated cases 
to the screen.

Before you run your simulation, we will "set the seed". We all will set the seed
to 100. This means that everyone's simulation will yield the exact same dataset.
```{r set-seed}
set.seed(100) 
# execute this line before you write your simulation code.
# only execute the set.seed() function one time. 
```

```{r simulate-200-tests}
sim_01 <- "YOUR ANSWER HERE"
sim_01
check_problem2()
```

Notice that `sim_01` is not a data frame, rather it is a vector of numbers. The 
following code stores `sim_01` as a data frame and changes its variable name. 
Run this code and view `sim_01` in the Viewer pane.

```{r, message=F, warning=F}
library(dplyr)
sim_01 <- as.data.frame(sim_01) # watch what happens to sim_01 in your environment
names(sim_01) # prints the variable names in the sim_01 data frame
sim_01 <- sim_01 %>% rename(covid_19 = sim_01)
```

\newpage

3. [2 points] Write code to determine the number of COVID-19 cases that occurred in your 
simulation, and the corresponding proportion with COVID-19. Assign your 
output the name `output_01`. Print `output_01` to the screen. Hint: Use `dplyr` functions
to do this.

```{r, problem-3}
output_01 <- "YOUR ANSWER HERE"
output_01
check_problem3()
```

\newpage

4. [2 points] Re-run your simulation four more times and assign the output to a unique 
name each time. Print to the screen the number and proportion after each run. 
(Basically, "recycle" the code above four times)

```{r, problem-4}
sim_02 <- "YOUR ANSWER HERE"
output_02 <- "YOUR ANSWER HERE"
output_02
sim_03 <- "YOUR ANSWER HERE"
output_03 <- "YOUR ANSWER HERE"
output_03
sim_04 <- "YOUR ANSWER HERE"
output_04 <- "YOUR ANSWER HERE"
output_04
sim_05 <- "YOUR ANSWER HERE"
output_05 <- "YOUR ANSWER HERE"
output_05
check_problem4()
```

\newpage

5. [1 point] Suppose that rather than simulating 5 samples of size 200, we simulated 5 samples
of size 1000. In 1-2 sentences, how would you expect the group of proportion 
estimates from part e) to be different? Comment both on the accuracy of these 
values at predicting the true value, and their variance. If you're not sure, you
can re-run your simulation with a larger sample size and see how the results 
change to deduct the difference.


[TODO: YOUR ANSWER HERE]


\newpage

**[8 points] Part 2: Probability of Cancer and Smoking Status**

Approximately 1.7 million Americans have cancer and 34.2 million Americans smoke cigarettes. The number of individuals who both have cancer and smoke is 300,000. Among individuals with cancer, approximately 40% smoke cigarettes. The total US population was approximately 328 million.

references for these stats:

* https://gis.cdc.gov/Cancer/USCS/DataViz.html
* https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm#:~:text=In%202018%2C%20nearly%2014%20of,with%20a%20smoking%2Drelated%20disease.
* https://www.cdc.gov/media/releases/2016/p1110-vital-signs-cancer-tobacco.html#:~:text=Each%20year%20between%202009%20and,but%20progress%20has%20been%20made.

\newpage

6. [2 points] Calculate the probability that a randomly chosen American will have cancer. 
Calculate the probability that a randomly chosen American will be a smoker. Convert
to percentages and round to two decimal places. Save these values as the vector p6 with the proportion for cancer first then smokers. Do not include the %.

```{r, problem-6}
p6 <- c("YOUR ANSWER HERE")
p6
check_problem6()
```



\newpage

7. [2 points] Without using the number of Americans who both have cancer and are smokers provided in the question,
calculate the probability that someone will have both cancer and a smoker status. Convert to a percent rounded to two decimal places and save to object p7. Don't include the percent in your answer.
```{r, problem-7}
p7 <- "YOUR ANSWER HERE"
p7
check_problem7()
```



\newpage

8. [2 points] Are cancer and smoker status independent? Show work to support your 
answer. Uncomment your selection.

```{r, problem-10}
#p8 <- "independent"
#p8 <- "not independent"
check_problem8()
```




\newpage

9. [2 points] In general, does P(A|B) equal P(B|A)? Calculate P(C | S) and
report whether or not it is equal to P(S | C).

[TODO: YOUR ANSWER HERE]



\newpage

**[9 points] Part 3: Screening for COVID-19**

Background reading: Read pages 258-261 (and optionally 261-264) of Baldi & 
Moore, Edition 4. (For earlier editions, look for the section on diagnostic 
testing in medicine or on screening which covers sensitivity, specificity, 
negative predictive value, and positive predictive value). 

*COVID-19 RT-PCR Test:*

The FDA approved emergency authorization for the COVID-19 RT-PCR test made by LabCorp.The RT-PCR test positively identifies SARS-CoV-2 95% of the time, and it gives a negative result for 100% of individuals who do not have COVID-19. Suppose 2% of the population actually has COVID-19.

\newpage

10. [3 points] Use probability notation to express the three probabilities cited above in the excerpt for COVID-19 RT-PCR Test. Make sure to  
define each event using a capital letter (or two). Provide the terminology for the 95% and 100% values based on your readings.

[TODO: YOUR ANSWER HERE]




\newpage



11. [3 points] What percent of COVID-19 tests in this target population would be positive? Answer
this question by using absolute frequencies. 
Show your work.

[TODO: YOUR ANSWER HERE]



\newpage


12. [1 point] Calculate the positive predictive value from question 11 based on your previous work from question 11. Store the answer as a percentage rounded to one decimal place in the object p12.


```{r}
p12 <- "YOUR ANSWER HERE"
p12
check_problem12()
```



\newpage
**13. [4 points] BONUS** At the end of every homework, you will have the opportunity to obtain up to *4 extra credit points* on that assignment. 

+1 point for developing a multiple choice question on the material in the homework

+1 point for developing a true/false question on the material 

+2 points for developing a free response question.

To get full credit, you need to write the question, write an answer, provide a rubric for grading the answer, and say something about why/how you feel this question addresses a concept in this homework. You can only write **one question per category**.

*Note: If you are including code as part of your questions, please include an additional code chunk that loads the required packages and the required dataset.*

[ YOUR ANSWER HERE ]

\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su20/hw/hw04; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.
