---
title: 'Lab 6: Introducing the Central Limit Theorem'
output: pdf_document
---

* Due date: Tuesday, July 28th 10:00pm.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab06.RAGS.R")
```

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage

--------------------------------------------------------------------------------

# (Optional) Relevant Textbook Exercises
The following questions in your textbook are helpful practice for 
understanding today's material:

Baldi and Moore: Ex. 13.5, 13.8, 13.9, 13. 10, 13.12, 13.14


# Introduction 

You will use a central data source (a Google sheet) attached to the lab to
develop a very concrete idea of sampling distributions, and to see the central 
limit theorem in action.  

# The underlying population

Suppose you had a data frame containing the **entire population** of all
residents of Alameda County. You have data on three variables:

1.  Born either out (=1) versus in (=0) the county.
2.  Number of siblings (integer)
3.  Number of visits to the hospital last year

1. Read in the data, L06_Alameda.csv, and assign it to the name `alameda_pop`. Calculate the true 
(population) mean, and make histograms or bar charts of the distribution for each 
variable.  

```{r readin, message = F, warning = F}
library(dplyr)
library(ggplot2)
library(readr)

#### YOUR CODE GOES HERE ####

```

Notice that the distribution of number of siblings and number of hospital visits
are discrete distributions and both skewed right. Today, we will focus on the 
distribution of the number of siblings `num_sibs`. Remember, we know the population
mean exactly, because we have all the data. You calculated the underlying population
mean in the code chunk above.

### Calculating the sampling distribution

We are now going to look at the approximate **sampling distribution for the sample mean**
of the `num_sibs` in (live) action. Remember from earlier lectures that a 
**sampling distribution** is a distribution for a statistic, like the sample 
proportion or the sample mean.

Each student in prior semesters was tasked during lab with repeatedly taking a random sample of the population. 
Once they took their sample they computed the sample mean and added it to a 
shared Google sheet linked below. As the data was added to the Google sheet, results were
graphed to illustrate how the sampling distribution varied for 
increasingly larger sample sizes.

The columns in the sheets are  `n (Sample_size)` `mean(numSibs)` `Name`.

### Your peers' task (which formulated the google sheet) 

1. Your peers randomly generated 10 simple random samples of size $n=5$ from the population from each of the three target populations and got the average of each of these to get 10 averages.  They then re-ran the random sampling procedure 5 times, getting the average each time and pasting it in the google doc. They pasted it into the provided google sheet making sure to type in the sample size the average was based upon in the appropriate column.

One sample has been provided to you. Try it yourself 10 times and record your results in the vector below.


```{r size5}
#### YOUR CODE GOES HERE ####

#repeat this code 10 times.
size_5 <- sample_n(alameda_pop, 5, replace = FALSE)
size_5 %>% summarise(mean_numSibs=mean(num_sibs))




#once finished, add your samples to this vector
#do not put them in quotations.
size_5_samples <- c("YOUR SAMPLES HERE")


size_5_samples

check_problem1()
```

After they calculated 10 sample means using the above code, they copy and pasted their
data into the google sheet for their lab section. The link to the past 
google sheet is:

https://docs.google.com/spreadsheets/d/130FOigRrcdzyMpxSmzxEcdx2TOthUeKZFBpfJgHpbWk/edit?usp=sharing

Look at the plot of the **sampling distribution** for
the mean number of siblings when $n=5$.

- What is the range of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]


- What is the shape of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]


- What is the mean of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]



Once the class had examined the sampling distribution when $n=5$, they repeated the 
same steps for n=50. 


2. Write the code to repeat the sampling process for a sample size of $n=50$. Repeat the sampling 10 times and store your values in the vector.

```{r size50}
#### YOUR CODE GOES HERE ####


size_50_samples <- c("YOUR VALUES HERE")


size_50_samples

check_problem2()
```

After they calculated their 10 sample means, they navigated to the google sheet from 
before, but switch to the sheet with $n=50$ (you can switch tabs in the bottom left).

Look at the plot to the right on the google sheet; now with $n=50$

- What is the range of the sampling distribution of the mean? How does it compare
to when $n=5$?

[TODO: YOUR ANSWER HERE]


- What is the shape of the sampling distribution of the mean? How does it compare to when $n=5$? 

[TODO: YOUR ANSWER HERE]


- What is the mean of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]


3. Write the code to repeat the sampling process for a sample size of $n=500$. Repeat the sampling 10 times and store your values in the vector.

```{r size500}
#### YOUR CODE GOES HERE ####

size_500_samples <- c("YOUR VALUES HERE")


size_500_samples
  
check_problem3()
```

- What is the range of the sampling distribution of the mean? How does it compare to when $n=5$ and $n=50$?

[TODO: YOUR ANSWER HERE]


- What is the shape of the sampling distribution of the mean? How does it compare
to when $n=5$ and $n=50$?

[TODO: YOUR ANSWER HERE]


- What is the mean of the sampling distribution of the mean?

[TODO: YOUR ANSWER HERE]



Suppose you have 500 classmates, and they have done this lab and added their data to this 
sheet: https://docs.google.com/spreadsheets/d/1AXStOd618raoWvrBequxOh5CDwgisFJoHo50fmcKb_E/edit?usp=sharing

Open the link, and look at the resulting sampling distributions for $n=5$, $n=50$, and $n=500$.
This is what happens when you repeat the sampling 5,000 times.


4. For which sample size should the sampling distribution of the mean be most normal?
  
  (a) n=5
  (b) n=50
  (c) n=500

Assign your letter choice as a string. Example: sampleSize_answer<-"b"
```{r}
sampleSize_answer<- "REPLACE WITH a,b, or c. Keep the quotes"
sampleSize_answer


check_problem4()

```

\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su20/lab/lab06; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.

