---
title: 'Lab 4: Screening and Precision Public Health'
author: "Your name and student ID"
date: "today's date"
output: pdf_document
---
* Due date: Monday, July 20 10:00pm

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab04.RAGS.R")

```

# Screening

We will talk about a very important use of conditional probability in public 
health and medicine, which is the idea of tools that screen for health outcomes.
There are many examples of this, including mammograms for detection of breast 
cancer, the prostate specific antigen (PSA) for detection of prostate cancer, as 
well as tests for exposure to infectious diseases, and so forth.  We will 
consider two types of events: i) whether the subject truly has the health condition
of interest (let $D$ denote the disease of interest, and $D'$ its complement), 
and ii) whether a test was positive or not for this outcome (let $+$ denote a 
positive test and $-$ denote its complement, a negative test). There are several 
statistics that are used to evaluate the performance of a test, some of which 
are derivable from each other:

- Sensitivity: $P(+ \mid D)$ or the probability of test being positive if one 
has the disease.
- Specificity: $P(- \mid D')$ or the probability of test being negative given 
one does not have the disease. 
- Positive predictive value (PPV): $P(D \mid +)$ or the probability of having the 
disease if an individual tests positive.
- Negative predictive value (NPV):  $P(D' \mid -)$ or the probability of not 
having the disease if an individual tests negative.

Consider the following situation: Assume the total number of subjects is 10,000
and that, $P(D) = 0.05$, $P(+ \mid D) = 0.95$, $P(- \mid D') = 0.95$. This set up
implies that the disease is rare, but that a very accurate test exists (i.e., 
equally high sensitivity and specificity).

**1. Fill in the following  two-way table with the absolute frequencies using the 
information provided in the problem. (You can fill it in simply by typing the 
frequencies in the nine empty cells.):

|       | $D$ | $D'$ |  Total|
|-------|-----|------|-------|
|+      |     |      |       |
|-      |     |      |       |
| Total |     |      |       |

Create a vector, p1, with the entries of this table, entering one column at 
a time.  For example, if my table was as follows:

|       | $D$ | $D'$ |  Total|
|-------|-----|------|-------|
|+      |  1  |  2   |  3    |
|-      |  4  |  5   |  6    |
| Total |  7  |  8   |  9    |

Then I would enter my vector as:
```{r}
p1 <- c(1, 4, 7, 2, 5, 8, 3, 6, 9)
```

```{r}
p1 <- "<<<<YOUR CODE HERE>>>>"
p1


check_problem1()
```

\newpage


**2. Calculate the PPV using the entries from your table. 
Store the PPV as a percentage.


```{r}
p2 <- "<<<<YOUR ANSWER HERE>>>>"
  

check_problem2()

```

\newpage
  
**3. Re-do the two-way table and re-calculate the PPV, this time assuming that 
$P(D)=0.02$. (Note that $P(+ \mid D) = 0.95$, $P(- \mid D') = 0.95$, as with 
the previous question.) 

|       | $D$ | $D'$ |  Total|
|-------|-----|------|-------|
|+      |     |      |       |
|-      |     |      |       |
| Total |     |      |       |


Similarly, create a vector, p3, with the entries of this table, entering one column at 
a time.
  
```{r}
p3 <- "<<<<YOUR CODE HERE>>>>"
p3


check_problem3()

```
 
\newpage 
    
**4. Again, calculate the PPV using the entries from your table. 
Store the PPV as a percentage rounded to 1 decimal place.
  
```{r}
p4 <- "<<<<YOUR ANSWER HERE>>>>"
  

check_problem4()

```
  

\newpage

**5. Explain why the sensitivity is so high, but the PPV is low for the first 
calculation and even lower for the second calculation.


[TODO: YOUR ANSWER HERE]
  


  
\newpage
  
# Precision public health


One of the goals of public health research is to group people by risk factors or
demographic variables so that decision-makers can predict, with actionable 
accuracy, which groups are at high and low risk of an adverse health outcome. In
this set of questions, we consider stratified two-way tables, which are two-way
tables specific to levels of a third grouping variable. Here, the adverse health
outcome is coronary heart disease (CHD), which we represent by $D$.  We study 
two categorical risk factors, smoking (defined by $S$ for smoking and $S'$ for 
no smoking) and age (defined by $A$ for older age and $A'$ for younger age). 

First,read in the aggregated data set. The last column (n) is the number of 
individuals in each group. 
                       
```{r makepop, message=F,warning=F}
library(dplyr)
library(tidyverse)
chd_dat <- read_csv("CHD.csv")

chd_dat
```

From this table:
                       
**6. Calculate the following probabilities. Convert your answers to percentages at the nearest whole number. Input the values (without the %) in the order below into the vector p6**: 
                       
- $P(D \mid A', S)$
- $P(D \mid A', S')$ 
- $P(D \mid A, S)$ 
- $P(D \mid A, S')$  

```{r}
p6 <- c("YOUR ANSWER HERE")
p6


check_problem6()
```



       
If you prefer, you can do these calculations by hand based on `chd_dat`. Some 
students might wish to use R commands to calculate these probabilities. There 
are **many** ways to do this. You could use `dplyr` functions to perform the 
calculations. Alternatively, here are some new functions for those of you 
interested in learning more R. (Note that these new functions won't be tested, 
they are for your information only, and to expose you to more of the R language!). 
You could consider using the `uncount()` function to expand the data based upon 
the numbers in each group (i.e., `n`) and assign the expanded data frame to a 
new name.Then, you can use the `tabyl` function from the `janitor` package to 
create stratified two-way tables, and the relevant `adorn_` functions from 
`janitor` to convert the frequencies to percentages.
  
```{r, message=F,warning=F}
```  


 
 
\newpage

**7. What do these numbers imply about smoking and the risk of CHD?
 
 
[TODO: YOUR ANSWER HERE]



\newpage

**8. Calculate the marginal probability of CHD. This can be written as $P(D)$. Store it in the variable p8.

```{r}
p8 <- "YOUR ANSWER HERE"
p8


check_problem8()
```


```{r, message=F,warning=F}
```




\newpage

**9. Calculate the conditional probabilities The $P(D \mid A')$ and $P(D \mid A)$. Store them in the vector p9 rounded to the nearest whole number percentage.

```{r}
p9 <- c("YOUR ANSWER HERE") 
p9


check_problem9()
```






\newpage

**10. If you had an intervention that could eliminate the risk of smoking on CHD, 
which group (defined by age) would see the biggest change in CHD from this 
intervention?
  
[TODO: YOUR ANSWER HERE]
  




\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su20/lab/lab04; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.
