---
title: "Homework 3"
author: "Your name and Student ID"
date: "Today's date"
output: pdf_document
---

* Due date: Thursday, July 16 10:00pm.
* Remember: autograder is meant as sanity check ONLY. It will not tell you if you have the correct answer. It will tell you if you are in the ball park of the answer so *CHECK YOUR WORK*.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/hw03.RAGS.R")
```

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage

--------------------------------------------------------------------------------

```{r load-libraries, message=F}
library(readr)
library(dplyr)
library(ggplot2)
library(broom)
library(forcats)
```

### Predicting total blood cholesterol levels by age and BMI

**Problem**: According to the Centers for Disease Control, high blood cholesterol raises the risk for heart disease, the leading cause of death in the United States, and for stroke, the fifth leading cause of death. Total cholesterol levels less than 200 mg/dL are considered desirable for adults. However, a 2017 report by the American Heart Association showed that 95 million U.S. adults age 20 or over have total cholesterol levels higher than 200 mg/dL.

**Your Task**: You are tasked with determining how total cholesterol levels are associated with age, for female patients who have a body mass index (bmi) in the "normal" range (18.5 to 24.9).

**Plan**: You have chosen to use tools to examine relationships between two variables to address the problem. In particular, scatter plots and simple linear regression.

**Data**: You have access to the dataset `framingham.csv`, a dataset from the Framingham Heart Study, a long-term, multigenerational cardiovascular cohort study of residents in Framingham, Massachusetts.

**Analysis and Conclusion**: In this assignment you will perform the analysis and make a conclusion to help answer the problem statement.

\newpage

**1. [1 point] Please type one line of code below to import these data into R. Assign the data to `fram_data`. Execute the code by hitting the green arrow and ensure the data set has been saved by looking at the environment tab and viewing the data set by clicking the table icon to the right of its name.**

```{r load-insure-data}
fram_data <- "<<<<YOUR CODE HERE>>>>"

# ------- REMINDER -------
# The checks on this homework are only provided as sanity checks.
# They do not guarantee correctness.
# ------------------------

check_problem1()
```

\newpage

**Use the space below to use the functions from lecture to get to know your dataset. Execute these functions line by line so you can look at their output, and examine these data.**

```{r initial-looks}
dim(fram_data)
names(fram_data)
str(fram_data)
head(fram_data)
```

\newpage

**2. [1 point] How many individuals are in the dataset? Assign this number to p2.**
```{r p2}
p2 <- "<<<<YOUR CODE HERE>>>>"


check_problem2()
```

\newpage

**3. [1 point] What are the nominal variable(s) in the dataset? Assign the names of these variable(s) to a vector of strings, p3.**
```{r p3}
#Hint: If I wanted p0 to be a vector of strings of fruits, I could
#do something like:
#p0 <- c("apple", "orange", "banana")

p3 <- "<<<<YOUR CODE HERE>>>>"


check_problem3()
```

\newpage

**4. [1 point] What are the ordinal variable(s) are in the dataset? Assign the names of these variable(s) to a vector of strings, p4.**
```{r p4}
p4 <- "<<<<YOUR CODE HERE>>>>"


check_problem4()
```

\newpage

**5. [1 point] What are the continuous variable(s) are in the dataset? Assign the names of these variable(s) to a vector of strings, p5.**
```{r p5}
p5 <- "<<<<YOUR CODE HERE>>>>"


check_problem5()
```

\newpage

**6. [1 point] What variable in the dataset could be considered discrete? Assign the name of this variable to a vector of strings, p6.**
```{r p6}
p6 <- "<<<<YOUR CODE HERE>>>>"


check_problem6()
```

\newpage

**Run the following code by removing the "#" symbol in front of each of the six lines and executing the code chunk. Remind yourself what the `mutate()` function does in general, and notice that a new function `case_when()` is also being used.**

```{r mutate-case-when}
fram_data <- fram_data %>%
  mutate(bmi_cat = case_when(bmi < 18.5 ~ "Underweight",
                             bmi >= 18.5 & bmi < 25 ~ "Normal",
                             bmi >= 25 & bmi < 30 ~ "Overweight",
                             bmi >= 30 ~ "Obese")
                  )
```

**7. [1 point] What did the above code achieve?:**

[TODO: YOUR ANSWER HERE]


\newpage

**8. [1 point] What type of variable is `bmi_cat`? Uncomment one of the choices below.**
```{r p8}
# p8 <- "ordinal"
# p8 <- "nominal"
# p8 <- "continuous"
# p8 <- "discrete"

# This only checks that you've selected an answer, not its correctness.
check_problem8()
```

\newpage

**9. [1 point] Read the Problem and Your Task statements proposed at the beginning of this exercise. Who belongs to the population of interest? Uncomment one of the choices below.** 
```{r p9}
# p9 <- "Females of normal BMI"
# p9 <- "Females of overweight BMI"
# p9 <- "Females who have abnormal BMI"
# p9 <- "All people at risk of high total cholesterol"

# This only checks that you've selected an answer, not its correctness.
check_problem9()
```


**10. [1 point] Using a dplyr function, make a new dataset called `fram_subset` containing the population of interest.**

```{r dplyr-the-data}
fram_subset <- "<<<<YOUR CODE HERE>>>>"


check_problem10()
```

\newpage

**11. [3 points] Make a scatter plot of the relationship between `age` and total cholesterol levels `totChol` for the population of interest. Give your plot an informative title.**

```{r scatter-plot}
p11 <- "<<<<YOUR CODE HERE>>>>"
p11


check_problem11()
```

\newpage

**12. [2 points] Run a linear regression model on the relationship between `age` and `totChol`. Think about which variable is explanatory (X) and which is response (Y). Assign the regression model to the name `fram_model`. Then type `tidy(fram_model)` below the model's code and execute both lines.**

```{r regression-mod}
fram_model <- "<<<<YOUR CODE HERE>>>>"
# <<<<YOUR CODE HERE>>>>


check_problem12()

```

\newpage

**13a. [1 point] Interpret the slope parameter:**

[TODO: YOUR ANSWER HERE]


**13b. [1 point] Interpret the intercept parameter:**

[TODO: YOUR ANSWER HERE]



**13c. [1 point] Does the intercept make sense in this context?:**

[TODO: YOUR ANSWER HERE]

\newpage

**14. [1 point] Add the line of best fit to your scatterplot by copying and pasting the plot's code from Problem 11 into the chunk below and adding a `geom` that can be used to add a regression line:**

```{r scatter-plot-with-line}
p14 <- "<<<<YOUR CODE HERE>>>>"
p14


check_problem14()
```

\newpage

**15. [1 point] Calculate the r-squared value for `fram_model` using a function learned in class. Assign this value to fram_model_r2.**

```{r rsquared}
fram_model_r2 <- "<<<<YOUR CODE HERE>>>>"


check_problem15()
```

\newpage

**16. [2 points] Calculate the correlation between `age` and `totChol` using the subset `fram_subset`. Also calculate correlation squared. You should use `summarize()` and name the two new columns `corr` and `corr_sq`. What do you notice about the relationship between the correlation and r-squared values that you calculated earlier?**

```{r correlations}
p16 <- "<<<<YOUR CODE HERE>>>>"


check_problem16()
```

\newpage

### PART B

**Your supervisor asks you to extend your analysis to consider other females with BMIs classified as overweight or obese. In particular, she wanted to know if the relationship between age and total cholesterol levels is different for different BMI groups. You can use data visualization coupled with your skills in linear regression to help answer this question.**

**17. [1 point] Make a new dataset called `fram_females` that includes females with BMIs classified as normal, overweight, or obese.**

```{r dplyr-the-data-smokers}
fram_females <- "<<<<YOUR CODE HERE>>>>"
  

check_problem17()
```

\newpage

**18. [1 point] Make a scatter plot that examines the relationship between `age` and `totChol` separately for normal, overweight, and obese females. A `facet_` command may help you.**

```{r ggplot-by-bmi}
p18 <- "<<<<YOUR CODE HERE>>>>"
p18


check_problem18()
```

**Is there something out of order with your plot you just made? The issue is that the plot is automatically displayed by listing the BMI categories alphabetically. Uncomment and run the following code chunk to assign an ordering to the values of `bmi_cat`.**: 

```{r fct-relevel}
fram_females <- fram_females %>%
mutate(bmi_cat_ordered = forcats::fct_relevel(bmi_cat, "Normal", "Overweight", "Obese"))
```

\newpage

**19. [1 point] Re-run your plot code, but this time, facet using `bmi_cat_ordered`.**

```{r ggplot-by-bmi-2}
p19 <- "<<<<YOUR CODE HERE>>>>"
p19


check_problem19()
```

\newpage

**20. [3 points] Run a separate linear model for each BMI group. To do this, you will need to subset your data into the three groups of interest first. Call your models `normal_mod`, `overweight_mod`, and `obese_mod`. Use the `tidy()` function to display the output from each model.**

```{r more-models-more-outputs}

#subset your data here:
# "<<<<YOUR CODE HERE>>>>"
# "<<<<YOUR CODE HERE>>>>"
# "<<<<YOUR CODE HERE>>>>"

#create your linear models here:
normal_mod <- "<<<<YOUR CODE HERE>>>>"
overweight_mod <- "<<<<YOUR CODE HERE>>>>"
obese_mod <- "<<<<YOUR CODE HERE>>>>"

#tidy your model output here:
# "<<<<YOUR CODE HERE>>>>"
# "<<<<YOUR CODE HERE>>>>"
# "<<<<YOUR CODE HERE>>>>"


check_problem20()
```

\newpage

**For the next three problems, use the models to predict the total cholesterol level for a 35-year old by weight category. You don't need an R function to make these predictions, just the output from the model. Show your work for each calculation by writing the mathematical expression in and round the cholesterol level to 2 decimal levels.**

**21. [1 point] ...among normal BMI group:**
```{r predict-1}
p21 <- "<<<<YOUR CODE HERE>>>>"


check_problem21()
```


**22. [1 point] ...among overweight BMI group:**
```{r predict-2}
p22 <- "<<<<YOUR CODE HERE>>>>"


check_problem22()
```

**23. [1 point] ...among obese BMI group:**
```{r predict-3}
p23 <- "<<<<YOUR CODE HERE>>>>"


check_problem23()
```

\newpage

**24. [3 points] In three sentences maximum, (1) comment on the direction of the association, (2) comment on how much the slopes vary across the BMI groups, and (3) how much the prediction for a 35-year old varies.**

[TODO: YOUR ANSWER HERE]


\newpage

**25. [4 points] BONUS** At the end of every homework, you will have the opportunity to obtain up to *4 extra credit points* on that assignment. 

+1 point for developing a multiple choice question on the material in the homework

+1 point for developing a true/false question on the material 

+2 points for developing a free response question.

To get full credit, you need to write the question, write an answer, provide a rubric for grading the answer, and say something about why/how you feel this question addresses a concept in this homework. You can only write **one question per category**.

*Note: If you are including code as part of your questions, please include an additional code chunk that loads the required packages and the required dataset.*

[ YOUR ANSWER HERE ]


\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su20/hw/hw03; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.
